---
title: "RPMI_YPD combined Plate Reader analysis"
author: "Liz Hughes"
date: "23/11/2022"
output:
  html_document:
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , warning=FALSE, message=FALSE)
```


This analysis takes parts from Rmd files RPMI_PR_20200728.Rmd and nYPD_PR_20200807.Rmd to create a combined figure.
Test WT H99 and Gat 201 deletion mutants 003(Madhani) and 004(Bahn)in RPMI and YPD at 37 degrees in TECAN plate reader.
7 Bioreps, 3 Techreps each. (NB only 2 Techreps for Biorep 7)
Grow 5ml culture from colony (1 colony = 1 Biorep)in YPD, 30C, 250 rpm ON.
Seed at OD 600nm = 0.2 , 200 ul per well.
Run for 3 days.

```{r load_packages}
library(tidyverse)
library(reshape2)
library(cowplot)
library(extrafont)

```

```{r load_functions}
read_platereader_csv <- function(file, ...) {
  # read plate reader data in .csv format and convert time to hours, from seconds.
  readr::read_csv(file = file, ...) %>%
    dplyr::mutate(Time = Time/3600)
}

read_platemap_csv <- function(file, strain_df = NULL, ...) {
  platemap <- readr::read_csv(file = file, ...)
  if ( !is.null(strain_df) ) {
    platemap <- platemap %>%
      dplyr::left_join(strain_df, by = "Strain")
  }
  platemap
}

reshape_annotate_rawod <- function(rawod, platemap, 
                                   od.name = "OD595", 
                                   melt.id = "Time", 
                                   well.name = "Well") {
  rawod %>%
    # Tidy the data using the melt function from reshape2
    reshape2::melt(id = melt.id, 
                   variable.name = well.name,
                   value.name =  od.name) %>%
    # Annotate by joining with the platemap
    inner_join(platemap, by = well.name)
}

summarise_OD_fixed <- function(annotated_od, od.name = "OD595", ...) {
  # summarise OD fixed value across many times
  # here use `...` input to pass conditions to filter
  # e.g. summarise_OD_fixed(annotated_od, od.name = "OD595", is.na(Strain))
  annotated_od %>%
    dplyr::filter(...) %>%
    dplyr::summarise(OD_median = median(OD595),
                     OD_mean   = mean(OD595),
                     OD_max    = max(OD595),
                     OD_min    = min(OD595),
                     .groups = "keep")
}

summarise_OD_bytime <- function(annotated_od, od.name = "OD595", groups = "Time", ...) {
  # summarise OD at each time
  # here use `...` input to pass conditions to filter
  # e.g. summarise_OD_bytime(annotated_od, od.name = "OD595", groups = "Time", is.na(Strain))
  annotated_od %>%
    dplyr::filter(...) %>%
    dplyr::group_by(Time) %>%
    dplyr::summarise(OD_median_time = median(OD595),
                     OD_mean_time   = mean(OD595),
                     OD_max_time    = max(OD595),
                     OD_min_time    = min(OD595),
                     .groups = "keep")
}

normalise_OD_2ways <- function(annotatedod, ...) {
  # normalise OD across a plate
  # will NOT work if there are multiple media in a plate; that would require an extra group_by or filtering internally.
  # here use `...` input to pass conditions to filter
  # e.g. normalise_OD(annotated_od, is.na(Strain))
  # e.g. normalise_OD(annotated_od, Well %in% c("A1","A2","A3")) 
  
  normvalues_fixed <- 
    summarise_OD_fixed(annotatedod, od.name = "OD595", ...)
  normvalues_bytime <- 
    summarise_OD_bytime(annotatedod, od.name = "OD595", groups = "Time", ...)
  annotatedod %>%
    mutate(OD_corrected_fixed = OD595 - normvalues_fixed$OD_median) %>%
    left_join(normvalues_bytime, by = "Time") %>% 
    mutate(OD_corrected_time = OD595 - OD_median_time) 
}
```

```{r shared_formatting}
theme_set(
  theme_cowplot(font_size = 12, font_family = "Arial",
                rel_large = 1) + 
    theme(strip.background = element_blank(),
          panel.border = 
            element_rect(color = "grey90", 
                         fill = NA, 
                         linetype = 1, 
                         size = 1),
          panel.grid.major.y = 
            element_line(size = 0.5, 
                         colour = "grey90"))
)

scale_time_hrs <- 
  scale_x_continuous(name = "Time (hours)", 
                     breaks = c(0, 12, 24, 48, 72),
                     limits = c(0, 72),
                     expand = c(0, 0))
```


# Set up data frame of nice strain names, colours, styles.

```{r nice_names}
strain_df <- tibble( Strain = c("H99", "Gat201(B)", "Gat201(M)", "C26", "C30" ),
                     Strain_nice = c("GAT201", "gat201∆b","gat201∆m", "GAT201-C1", "GAT201-C2"),
                     colour = c("grey20", "magenta3", "darkred", "blue3", "cyan4"))

scale_color_strain <- 
  scale_color_manual("Strain",
                     breaks = strain_df$Strain_nice,
                     values = strain_df$colour)
```



# Read in transposed data as csv file RPMI


```{r load_rawod_RPMI}
rawod_RPMI <- 
  read_platereader_csv("../Input/20200728_EH_PR2_RPMI_WT_Gat201TRSP.csv")
```

# Read in the Plate map data from csv file RPMI

Also combine with nice strain names for display.

```{r load_platemap_RPMI}
platemap_RPMI <- 
  read_platemap_csv("../Input/20200728-PR2_SetupCSV.csv",
                    strain_df = strain_df) 

head(platemap_RPMI, n=10)

```

# Reshape data and combine with the plate map, pairing them by Well RPMI

```{r reshape_annotate_RPMI}
annotated_RPMI <- 
  reshape_annotate_rawod(rawod = rawod_RPMI, platemap = platemap_RPMI)
```

## Plot blank ODs RPMI

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs_RPMI, fig.width = 5, fig.height = 4}
ggplot(data=filter(annotated_RPMI, is.na(Strain)),
       aes(x=Time, y=OD595,  group = Well)) + 
  geom_line() + 
  scale_time_hrs + 
  scale_y_continuous(limits = c(0, 0.2),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15)) +
  labs(title = "Blank wells RPMI", x="Time (hours)", y="Absorbance at 595 nm")
```

## Calculate median OD for blank wells for each growth medium RPMI

Remove well H2

```{r calc_blank_medians_RPMI}
# filter out columns 1 and 12 as well?
blank_OD_summary_RPMI <- 
  summarise_OD_fixed(annotated_RPMI, od.name = "OD595",
                     is.na(Strain),
                     Well != "H2") 

print(blank_OD_summary_RPMI)

blank_OD_time_RPMI <- 
  summarise_OD_bytime(annotated_RPMI, od.name = "OD595",
                     is.na(Strain),
                     Well != "H2") 

print(blank_OD_time_RPMI)
```

# Subtract blank OD to make corrected OD and Plot OD_corrected v Time (hrs) RPMI

```{r subtract_blank_OD_RPMI}
normalisedOD_RPMI <- 
  normalise_OD_2ways(annotated_RPMI, is.na(Strain))
```

## Plot OD all stat_summary(geom -"line") Highlight summary line RPMI

```{r plot_all_stat_summary_RPMI, fig.width = 5, fig.height = 4}
RPMI_plot <- ggplot(data = normalisedOD_RPMI %>%
                      filter(!is.na(Strain)), 
                    aes(x = Time, y = OD_corrected_time, color = Strain_nice)) + 
  geom_line(aes(group = Well), size = 0.2, alpha = 0.2) + 
  stat_summary(fun = "median", geom = "line", size = 1) +
  scale_time_hrs + 
  # scale_y_continuous(limits = c(0, 0.9), 
  #                    expand = c(0, 0), 
  #                    breaks = c(0, 0.25, 0.5, 0.75)) +
  scale_y_continuous(limits = c(0, 0.1),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1)) +
  labs(y = "OD595",
       title = "RPMI") +
  scale_color_strain

RPMI_plot
```


```{r save_png_RPMI}
ggsave("../Results/20200728_EH_PR2_RPMI_WT_Gat201TRSP_summaryplot.png",
       width = 5.1, height = 5)
```


# Read in transposed data as csv file YPD


```{r load_rawod_YPD}
rawod_YPD <- 
  read_platereader_csv("../Input/20200807_EH_PR4_YPD_WTGat201_TRSP.csv")
```

# Read in the Plate map data from csv file YPD

Also combine with nice strain names for display.

```{r load_platemap_YPD}
platemap_YPD <- 
  read_platemap_csv("../Input/20200807-PR4_SetupCSV.csv",
                    strain_df = strain_df) 

head(platemap_YPD, n=10)

```

# Reshape data and combine with the plate map, pairing them by Well YPD

```{r reshape_annotate_YPD}
annotated_YPD <- 
  reshape_annotate_rawod(rawod = rawod_YPD, platemap = platemap_YPD)
```

## Plot blank ODs YPD

To find how stable the ODs of the blank wells are.

```{r plot_blank_ODs_YPD, fig.width = 5, fig.height = 3.5}
ggplot(data=filter(annotated_YPD, is.na(Strain)),
           aes(x=Time, y=OD595,  group = Well)) + 
  geom_line() + 
  scale_time_hrs + 
  scale_y_continuous(limits = c(0, 0.2),
                     expand = c(0, 0),
                     breaks = c(0, 0.05, 0.1, 0.15)) +
  labs(title = "Blank wells YPD", x="Time (hours)", y="Absorbance at 595 nm")
```

## Calculate median OD for blank wells for each growth medium YPD

```{r calc_blank_medians_YPD}
# filter out columns 1 and 12 as well?
blank_OD_summary_YPD <- 
  summarise_OD_fixed(annotated_YPD, od.name = "OD595",
                     is.na(Strain)) 

print(blank_OD_summary_YPD)

blank_OD_time_YPD <- 
  summarise_OD_bytime(annotated_YPD, od.name = "OD595",
                     is.na(Strain)) 

print(blank_OD_time_YPD)
```

# Subtract blank OD to make corrected OD and Plot OD_corrected v Time (hrs) YPD

```{r subtract_blank_OD_YPD}
normalisedOD_YPD <- 
  normalise_OD_2ways(annotated_YPD, is.na(Strain))
```

## Plot OD all stat_summary(geom -"line") Highlight summary line YPD

```{r plot_all_stat_summary_YPD, fig.width = 5, fig.height = 3.5}
YPD_plot <- ggplot(data = normalisedOD_YPD %>%
                      filter(!is.na(Strain)), 
                    aes(x = Time, y = OD_corrected_time, color = Strain_nice)) + 
  geom_line(aes(group = Well), size = 0.2, alpha = 0.2) + 
  stat_summary(fun = "median", geom = "line", size = 1) +
  scale_time_hrs + 
  scale_y_continuous(limits = c(0, 0.9),
                     expand = c(0, 0),
                     breaks = c(0, 0.1, 0.25, 0.5, 0.75)) +
  # scale_y_continuous(limits = c(0, 0.1),
  #                    expand = c(0, 0),
  #                    breaks = c(0, 0.05, 0.1)) +
  labs(y = "OD595",
       title = "YPD") +
  scale_color_strain

YPD_plot
```


```{r save_png_YPD}
ggsave("../Results/20200807_EH_PR4_YPD_WTGat201_TRSP_summaryplot.png",
       width = 5.1, height = 5)
```



# Draft Multipanel Figure

```{r multifig, fig.height = 4, fig.width = 3.75}

plot_grid(
  RPMI_plot,
  YPD_plot,
  ncol = 1)

```


```{r save_png}
ggsave(filename = "../results/fig_platereader.png", width = 3.75, height = 4)
```

```{r save_svg}
ggsave(filename = "../results/fig_platereader.svg", width = 3.75, height = 4)
```


























