---
title: "RPMI_YPD comnbined PR analysis"
author: "Liz Hughes"
date: "23/11/2022"
output:
  html_document:
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , warning=FALSE, message=FALSE)
```


This analysis takes parts from Rmd files RPMI_PR_20200728.Rmd and nYPD_PR_20200807.Rmd to create a combined figure.
Test WT H99 and Gat 201 deletion mutants 003(Madhani) and 004(Bahn)in RPMI and YPD at 37 degrees in TECAN plate reader.
7 Bioreps, 3 Techreps each. (NB only 2 Techreps for Biorep 7)
Grow 5ml culture from colony (1 colony = 1 Biorep)in YPD, 30C, 250 rpm ON.
Seed at OD 600nm = 0.2 , 200 ul per well.
Run for 3 days.

```{r load_packages}
library("reshape2")
library(dplyr)
library(ggplot2)
library(tidyverse)
library(markdown)
library(cowplot)
```


# Read in transposed data as csv file RPMI

```{r csv_file}

rawdata <- read.csv("../Input/20200728_EH_PR2_RPMI_WT_Gat201TRSP.csv")

```

# Change time in seconds to time in hours RPMI

```{r mutate_time}

 rawdata_hrs<-mutate(rawdata, Time = Time/86400)

```

# Tidy the data using the melt function from reshape2 RPMI

```{r reshape}
reshaped <- melt(rawdata_hrs, id=c("Time", "Temp"), 
                              variable.name="Well",
                              value.name="OD595")

summary(reshaped)
```

# Read in the Plate map data from csv file RPMI

```{r Load_platemap}

platemap <- read.csv("../Input/20200728-PR2_SetupCSV.csv")
head(platemap, n=10)


```

# Combine the reshaped data with the plate map, pairing them by Well RPMI

```{r join}

annotated <- inner_join(reshaped, platemap, by="Well")
 
```

## Calculate median OD for blank wells for each growth medium RPMI
Remove well H2

```{r calc_blank_medians}
blank_OD_summary <- annotated %>%
  filter(Strain=="") %>%
  filter(Well != "H2") %>%
  # filter out columns 1 and 12 as well?
  group_by(Medium) %>%
  summarise(OD_median=median(OD595),
            OD_mean=mean(OD595),
            OD_max=max(OD595),
            OD_min=min(OD595))
print(blank_OD_summary)
```

# Subtract blank OD to make corrected OD and Plot OD_corrected v Time (hrs) RPMI

```{r subtract_blank_OD}

normalisedOD <- annotated %>%
  left_join(blank_OD_summary, by="Medium") %>%
  mutate(OD_corrected = OD595 - OD_median)

```

## Plot OD all stat_summary(geom -"line") Highlight summary line RPMI

```{r plot_all_stat_summary}
ggplot(data=normalisedOD %>%
                     filter(Strain != ""), 
                      aes(x=Time, y=OD_corrected, color=Strain)) + 
       geom_line(aes(group = Well), size = 0.2, alpha = 0.2) + 
       stat_summary(fun = "median", geom = "line", size = 1) +
       scale_y_continuous(limits=c(0,0.1),expand=c(0,0), breaks = c(0, 0.05, 0.1)) +
  
       labs(x = "Number of Days",
            y = "Absorbance (595nm)",
            title = "Cell Culture Media") +
       theme_bw() +
       theme(text = element_text(size = 15))+
       theme(axis.title = element_text(size = 25)) +
  theme(axis.ticks.length = unit(0.3, "cm"))+
       theme(title = element_text(size = 30))+
       theme(legend.title = element_text(size = 25))+
  theme(panel.grid.minor = element_blank())+
       scale_color_brewer(palette="Set1")
```


# Read in transposed data as csv file YPD

```{r y_csv_file }
y_rawdata <- read.csv("../Input/20200807_EH_PR4_YPD_WTGat201_TRSP.csv")

```

# Change time in seconds to time in days YPD

```{r y_mutate_time}

 y_rawdata_hrs<-mutate(y_rawdata, Time = Time/86400)
```

# Tidy the data using the melt function from reshape2 YPD

```{r y_reshape}

y_reshaped <- melt(y_rawdata_hrs, id=c("Time", "Temp"), variable.name="Well",
                 value.name="OD595")
summary(y_reshaped)
```


# Read in the Plate map data from csv file YPD

```{r y_Load_platemap}
y_platemap <- read.csv("../Input/20200807-PR4_SetupCSV.csv")
head(y_platemap, n=10)

```

# Combine the reshaped data with the plate map, pairing them by Well YPD

```{r y_join}
y_annotated <- inner_join(y_reshaped, y_platemap, by="Well")

```

## Calculate median OD for blank wells YPD
Remove A1

```{r y_calc_blank_medians}
y_blank_OD_summary <- y_annotated %>%
  filter(Strain=="") %>%
  group_by(Medium) %>%
  summarise(OD_median=median(OD595),
            OD_mean=mean(OD595),
            OD_max=max(OD595),
            OD_min=min(OD595))
print(y_blank_OD_summary)
```

# Subtract blank OD to make corrected OD YPD

```{r y_subtract_blank_OD}

y_normalisedOD <- y_annotated %>%
  left_join(y_blank_OD_summary, by="Medium") %>%
  mutate(y_OD_corrected = OD595 - OD_median)

```

## Plot OD all stat_summary(geom -"line")

```{r y_plot_all_stat_summary}
ggplot(data=y_normalisedOD %>%
                     filter(Strain != ""), 
                      aes(x=Time, y=y_OD_corrected, color=Strain)) + 
       geom_line(aes(group = Well), size = 0.2, alpha = 0.2) + 
       stat_summary(fun = "median", geom = "line", size = 1) +
       scale_y_continuous(limits=c(0,0.9),expand=c(0,0)) +
       geom_hline(aes(yintercept=0.1), linetype="dashed", colour="black") +
       labs(x = "Number of Days",
            y = "Absorbance (595nm)",
            title = "Rich Culture Media") +
       theme_bw() +
       theme(text = element_text(size = 15))+
       theme(axis.title = element_text(size = 25))+
       theme(axis.ticks.length = unit(0.3, "cm"))+
       theme(title = element_text(size = 30))+
   theme(panel.grid.minor = element_blank())+
       theme(legend.title = element_text(size = 25))+
       scale_color_brewer(palette="Set1")
```




























