---
title: "YPD_PR_20200807"
author: "Liz Hughes"
date: "13/01/2022"
output:
  html_document:
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , warning=FALSE, message=FALSE)
```

#Plate reader assay4.

Test WT H99 and Gat 201 deletion mutants 003(Madhani) and 004(Bahn)in YPD.
7 Bioreps, 3 Techreps each. (NB only 2 Techreps for Biorep 7)
Grow 5mlculture from colony (1 colony = 1 Biorep)in YPD, 30C, 250 rpm ON.
Seed at OD 600nm = 0.2 , 200 ul per well.
Run for 3 days.
```{r load_packages}

library("reshape2")
library(dplyr)
library(ggplot2)
library(tidyverse)
library(markdown)
library(cowplot)
```



# Read in transposed data as csv file YPD

```{r y_csv_file }
y_rawdata <- read.csv("../Input/20200807_EH_PR4_YPD_WTGat201_TRSP.csv")

```

# Change time in seconds to time in days YPD

```{r y_mutate_time}

 y_rawdata_hrs<-mutate(y_rawdata, Time = Time/86400)
```

# Tidy the data using the melt function from reshape2 YPD

```{r y_reshape}

y_reshaped <- melt(y_rawdata_hrs, id=c("Time", "Temp"), variable.name="Well",
                 value.name="OD595")
summary(y_reshaped)
```


# Read in the Plate map data from csv file YPD

```{r y_Load_platemap}
y_platemap <- read.csv("../Input/20200807-PR4_SetupCSV.csv")
head(y_platemap, n=10)

```

# Combine the reshaped data with the plate map, pairing them by Well YPD

```{r y_join}
y_annotated <- inner_join(y_reshaped, y_platemap, by="Well")

```


# How stable are ODs of blank wells, to use for normalisation? YPD

```{r y_plot_blank_ODs}

ggplot(data=filter(y_annotated,Strain==""), 
      
       aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
 
       facet_grid(Medium ~ .) +
       labs(x="Number of Days", y="Absorbance at 595 nm")
```

Remove A1

## Calculate median OD for blank wells YPD

```{r y_calc_blank_medians}
y_blank_OD_summary <- y_annotated %>%
  filter(Strain=="") %>%
  group_by(Medium) %>%
  summarise(OD_median=median(OD595),
            OD_mean=mean(OD595),
            OD_max=max(OD595),
            OD_min=min(OD595))
print(y_blank_OD_summary)
```


# Subtract blank OD to make corrected OD YPD

```{r y_subtract_blank_OD}

y_normalisedOD <- y_annotated %>%
  left_join(y_blank_OD_summary, by="Medium") %>%
  mutate(y_OD_corrected = OD595 - OD_median)

```

## Plot OD all YPD

```{r y_plot_all}
ggplot(data=y_normalisedOD %>%
                      filter(Strain != ""), 
                      aes(x=Time, y=y_OD_corrected, color=Strain, group = Well)) + 
       geom_line() + 
       facet_grid(Medium ~ .) +
       scale_y_continuous(limits=c(0,1),expand=c(0,0))+
       geom_hline(aes(yintercept=0.5), linetype="dotted", colour="black") +
       labs(x = "Number of Days",
            y = "Absorbance (595nm)") +
      theme(text = element_text(size = 20))+
      theme_bw()+
      scale_color_brewer(palette="Set1")
```

## Plot OD all stat_summary(geom -"line")

```{r y_plot_all_stat_summary}
ggplot(data=y_normalisedOD %>%
                     filter(Strain != ""), 
                      aes(x=Time, y=y_OD_corrected, color=Strain)) + 
       geom_line(aes(group = Well), size = 0.2, alpha = 0.2) + 
       stat_summary(fun = "median", geom = "line", size = 1) +
       scale_y_continuous(limits=c(0,0.9),expand=c(0,0)) +
       geom_hline(aes(yintercept=0.1), linetype="dashed", colour="black") +
       labs(x = "Number of Days",
            y = "Absorbance (595nm)",
            title = "Rich Culture Media") +
       theme_bw() +
       theme(text = element_text(size = 15))+
       theme(axis.title = element_text(size = 25))+
       theme(axis.ticks.length = unit(0.3, "cm"))+
       theme(title = element_text(size = 30))+
   theme(panel.grid.minor = element_blank())+
       theme(legend.title = element_text(size = 25))+
       scale_color_brewer(palette="Set1")
```

```{r save_YPD_PR_Gat201_0.1.png}

ggsave("../Results/YPD_PR_Gat201_0.1.png",width = 7,height=5)
```

## Plot OD all stat_summary(geom -"line") remove excess lines YPD

```{r y_plot_all_stat_summary_nolines}

ggplot(data=y_normalisedOD %>%
                     filter(Strain != ""), 
                      aes(x=Time, y=y_OD_corrected, color=Strain)) + 
       #geom_line(aes(group = Well), size = 0.2, alpha = 0.2) + 
       stat_summary(fun = "median", geom = "line", size = 1) +
       scale_y_continuous(limits=c(0,0.9),expand=c(0,0)) +
       geom_hline(aes(yintercept=0.1), linetype="dashed", colour="black") +
       labs(x = "Number of Days",
            y = "Absorbance (595nm)",
            title = "Rich Culture Media") +
       theme_bw() +
       theme(text = element_text(size = 15))+
       theme(axis.title = element_text(size = 25))+
       theme(axis.title.y=element_blank()) +
       theme(axis.ticks.length = unit(0.3, "cm"))+
       theme(title = element_text(size = 30))+
       theme(panel.grid.minor = element_blank())+
       theme(legend.title = element_text(size = 25))+
      theme(legend.background = element_rect(fill = "transparent", color=NA))+
  theme(legend.key = element_rect(fill = "transparent", color=NA))+
       theme (plot.background =  element_rect(fill = "transparent", color=NA), 
         panel.background = element_rect(fill = "transparent", color=NA),)+
      theme(panel.border = element_rect(linetype = "solid", colour = "black", size = 2))+
       scale_color_brewer(palette="Set1")


```


```{r save_YPD_PR_Gat201_0.1_TP.png}

ggsave("../Results/YPD_PR_Gat201_0.1_TP.png",width = 6.6,height=5)
```




